Chapter 16: Particle-laden gravity currents: the lock-release slumping regime at the laboratory scale

### 2024-03-14 paper/main.tex

A two-fluid model, SedFoam \citep{chauchat2017}, is used to simulate lock-release configurations from set-ups 1 and 2. This model available for download at Zenodo \citep{bonamy2023} is implemented in the open-source computational fluid dynamics toolbox OpenFOAM and has been extensively validated for various sediment transport problems such as sheet-flow \citep{cheng2016,chauchat2022,mathieu2021,mathieu2022}, scour \citep{nagel2020,tsai2022}, ripples \citep{salimi2021a,salimi2021b}, avalanches \citep{montella2021} and granular collapse \citep{montella2023}. The simulations presented herein are 2D vertical and turbulence-averaged. The turbulence model is a $k-\epsilon$ model with additional two-phase flow terms for density stratification and turbulent drag dissipation. The closure for the particle-particle interactions is a frictional-collisional model based on a Coulomb rheology for the frictional part and the kinetic theory of granular flows for the collisional part. The two contributions are simply added as described in \cite{chauchat2017}. The drag force is modeled using the \cite{schiller1933} closure multiplied by a hindrance function of \cite{richardson1954}. In order to describe the influence of the slope, the total pressure is solved and the gravity vector is rotated while the mesh is kept aligned with the $x$ and $y$ axis.

@@

\subsection*{Acknowledgment}
We thank Jean-Dominique Barron (IMFT) and Sébastien Cazin (IMFT) for their support in carrying out the experiments on setup-up 1.
%
Sylvain Dauge is acknowledged for the full design of setup 2. Hervé Bellot and Brivaël Collin are thanked for their experimental support.
%
We also acknowledge Hugo Divel for his contribution to the preliminary lock-release simulations using SedFoam.
%
Finally, we would like to acknowledge the contributors of the open-source numerical Python libraries, including Matplotlib \citep{Hunter2007}, NumPy \citep{Harris2020} and SciPy \citep{Virtanen2020}, which provide an incredibly efficient ecosystem allowing scientific research in Python.

>>>>

A two-fluid model, SedFoam \citep{chauchat2017}, is used to simulate lock-release configurations from set-ups 1 and 2. This model available for download at Zenodo \citep{bonamy2023} is implemented in the open-source computational fluid dynamics toolbox OpenFOAM and has been extensively validated for various sediment transport problems such as sheet-flow \citep{cheng2016,chauchat2022,mathieu2021,mathieu2022}, scour \citep{nagel2020,tsai2022}, ripples \citep{salimi2021a,salimi2021b}, avalanches \citep{montella2021} and granular collapse \citep{montella2023}. The simulations presented herein are 2D Reynolds-averaged considering the streamwise and wall-normal direction. The turbulence model is a $k-\epsilon$ model with additional two-phase flow terms for density stratification and turbulent drag dissipation. The closure for the particle-particle interactions is a frictional-collisional model based on a Coulomb rheology for the frictional part and the kinetic theory of granular flows for the collisional part. The two contributions are simply added as described in \cite{chauchat2017}. The drag force is modeled using the \cite{schiller1933} closure multiplied by a hindrance function of \cite{richardson1954}. In order to describe the influence of the slope, the total pressure is solved and the gravity vector is rotated while the mesh is kept aligned with the $x$ and $y$ axis.

@@

\subsection*{Acknowledgment}
We thank Jean-Dominique Barron (IMFT) and Sébastien Cazin (IMFT) for their support in carrying out the experiments on setup-up 1.
%
Sylvain Dauge is acknowledged for the full design of setup 2. Hervé Bellot and Brivaël Collin are thanked for their experimental support.
%
We also acknowledge Hugo Divel for his contribution to the preliminary lock-release simulations using SedFoam.
%
We also acknowledge the contributors of the open-source numerical Python libraries, including Matplotlib \citep{Hunter2007}, NumPy \citep{Harris2020} and SciPy \citep{Virtanen2020}, which provide an incredibly efficient ecosystem allowing scientific research in Python.
%
Finally, we also would like to acknowledge the two anonymous reviewers for their comments and Prof. Marius Ungarish for his extensive suggestions.
